---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

### 1. Loading the data and check what the status


```{r, echo=TRUE}
activity <- read.csv('activity.csv')
head(activity)
```

```{r}
str(activity)
```


### 2. Loading the dplyr package 
```{r,message=FALSE}
library(dplyr)
```

### 3. Process/transform the data (if necessary) into a format suitable for your analysis
```{r}
dailysteptotal <-activity %>% group_by(date) %>% summarise(steps = sum(steps))

```



## What is mean total number of steps taken per day?

### 1. Calculate the total number of steps taken per day
```{r, echo=TRUE}
dailysteptotal <-activity %>% group_by(date) %>% summarise(steps = sum(steps))
head(dailysteptotal)
```

### 2. Make a histogram using base plotting system
```{r}
hist(dailysteptotal$steps,breaks = seq(0,25000,by = 2500),ylim = range(0,20,5),main = "Histogrm of Total Number of Steps Taken Each Day",xlab ="Total steps per day", col = "red")
```

### 3. Calculate and report the mean and median of the total number of steps taken per day
```{r}
meanD <- mean(dailysteptotal$steps,na.rm = T)
medianD <-median(dailysteptotal$steps,na.rm = T)
meanD
medianD
```

### Answer: The mean and median of the total number of steps taken per day is `r meanD` and `r medianD`.



## What is the average daily activity pattern?

### 1.Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
dailystepavg <-activity %>% group_by(interval) %>% summarise(steps = mean(steps,na.rm = T))
plot(x=dailystepavg$interval,y=dailystepavg$steps,type = "l",col="red",xlab = "Interval", ylab = "Steps")
```

### 2. Calculate Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps
```{r}
maxintervalstep <- subset(dailystepavg, steps == max(dailystepavg$steps))
a <- maxintervalstep$interval
```

### Answer: Interval at `r a` minutes contains the maximum number of steps.

## Imputing missing values

### 1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
missing <-sum(is.na(activity))

```

The total number of missing values is `r missing`.

### 2. Devise a strategy for filling in all of the missing values in the dataset. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
newact <- activity
for (i in 1:nrow(newact)) {
    if (is.na(newact$steps[i])) {
        newact$steps[i] <-  dailystepavg[which(newact$interval[i] == dailystepavg$interval), ]$steps
    }
}

head(newact)

sum(is.na(newact))
```

### 3.Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.

```{r}
NewstepT <-newact %>% group_by(date) %>% summarise(steps = sum(steps))

hist(NewstepT$steps,breaks = seq(0,25000,by = 2500),ylim = range(0,20,5),main = "Histogrm of Total Number of Steps Taken Each Day",xlab ="Total steps per day", col = "red")
```

```{r}
newmeanD <- mean(NewstepT$steps)
newmedianD <- median(NewstepT$steps)

newmeanD
newmedianD
```

### These formulas gives a mean and median of `r newmeanD` and `r newmedianD` respectively. The mean number is same but the median has a little change. Overall, there is no big impact on the analysis of data.



## Are there differences in activity patterns between weekdays and weekends?

### 1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
newact$date <-as.Date(newact$date,"%Y-%m-%d")
newact$weekday <-weekdays(newact$date)
newact$dayType <- ifelse(newact$weekday %in% c("Saturday","Sunday"), "weekends","weekdays")
head(newact)

```

### 2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.
```{r}
newstepavg <- aggregate(steps ~ interval + dayType, data=newact, mean)

head(newstepavg)
```

```{r}
library(lattice)

xyplot(steps ~interval | dayType, newstepavg,type = "l",layout = c(1,2),xlab="Interval", ylab = "Number of steps")

```


